Computational tools for representation theory of affine Lie algebras

### 2009-07-22 AffineLieAlgebras/ACMS2009/article.tex
The introduction of the article is rewritten in more scientific style, found several papers, which should be added to the bibliography.

In this article we will describe computation algorithms for representations of affine Lie algebras by example.

Theory of affine Lie algebras is based upon the theory of simple Lie
algebras. Affine Lie algebras are infinite-dimensional, but this
infinite-dimensionality is ``along one direction''.

So in computations we limit ourselves with some finite grade. 
Then the problem is to propose algorithms with lowest growths rate
with the increase of maximal grade.

@@

We consider integrable modules $L^{\mu}$ with the highest weight
$\mu$. As for simple Lie algebras the most computation-intensive task
is to find weight multiplicities (dimensions of weight subspaces). 
These multiplicities are usually encoded in form of string functions
and these functions are connected with generic theta-functions and
modular forms \cite{kac1990idl}. 
But it is easier to work with multiplicities for automated
computations.

>>>>

Since the affine Lie algebras are infinite-dimensional extensions of simple finite-dimensional Lie algebras, their representation theory is well understood \cite{kac1990idl}, \cite{wakimoto2001idl}, \cite{Wakimoto-Second-Book}. But infinite dimensionality poses a challenge for computational methods.

In this paper we discuss the tasks of construction of representations which is reduced to computation of weight multiplicities and of branching of representation of affine Lie algebra to representation of affine subalgebra.

The first task is usually solved using the algorithms based on the recursive Freudenthal formula for weight multiplicities \eqref{eq:1}. The summation is optimized by considering of Weyl symmetry \cite{kass1990ala}, \cite{Moody-Patera}. While these algorithms are reasonably fast, the use of Racah-like recurrent relation \eqref{eq:2} is theoretically more efficient \cite{Nazarov2008}. In practice speed depends upon the realisation. So we describe this alternative approach in section \ref{sec:representations} and show results of practical speed comparisons.

The task of calculation of branching coefficients is usually solved by the reduction to finite-dimensional case \cite{difrancesco1997cft}. In some physical applications this task is greatly simplified by the additional physical constraints, but for general case this approach is rather inconvenient since it is not easily formalisable for machine compuatation. We propose another approach based on recurrent relations for anomalous branching coefficients \cite{ilyin812pbc} in section \ref{sec:branching-rules} and discuss the method to optimize this computation. Then we describe our realisation of mentioned algorithms \ref{sec:details-implementation} and some applications \ref{sec:applications}.

@@

We consider integrable modules $L^{\mu}$ with the highest weight
$\mu$. As for simple Lie algebras the most computation-intensive task
is to find weight multiplicities (dimensions of weight subspaces). 
These multiplicities are usually encoded in form of string functions
and these functions are connected with generic theta-functions and
modular forms \cite{kac1990idl}.

It is possible to write general relations for string functions in order to express them as the combination of generic theta-functions \cite{Lyakhovsky-Unwritten-String-equations-with-fans}. But in these paper we limit ourselves to automated computation of weight multiplicities.
